---
title: "Week5_class"
output: html_document
requires: fpp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp)
```


```{r Stationarity}
par(mfrow = c(1, 3))
plot(dj, main = "(a)", xlab = "Day")
plot(diff(dj), main = "(b)", xlab = "Day")
plot(strikes, main = "(c)", xlab = "Year")
plot(hsales, main = "(d)", xlab = "Year")
plot(eggs, main = "(e)", xlab = "Year")
plot(pigs, main = "(f)", xlab = "Year")
plot(lynx, main = "(g)", xlab = "Year")
plot(beer, main = "(h)", xlab = "Year")
plot(elec, main = "(i)", xlab = "Year")

par(mfrow = c(1, 2))
Acf(dj)
Acf(diff(dj))

Box.test(diff(dj), type = "L", lag = 10)

par(mfrow = c(1, 3))
plot(log(usmelec), ylab="Logs")
plot(diff(log(usmelec),12), ylab="
     Differenced logs")
plot(diff(diff(log(usmelec),12),1), ylab="Doubly differenced logs")


par(mfrow=c(1,1))

```

Creating AR(1) and AR(2) charts
```{r AR}
set.seed(1)
par(mfrow = c(1, 2))
testAR1 <- 10 + arima.sim(list(ar = -0.8), n= 100)
plot(testAR1, 
     ylab = "", main = "AR(1)")
testAR2 <- 20 + arima.sim(list(ar = c(1.3, -0.7)), n = 100)
plot(testAR2, 
     ylab = "", main = "AR(2)")

```

Creating MA(1) and MA(2) charts
```{r MA}
set.seed(2)
par(mfrow = c(1, 2))
testMA1 <- 20 + arima.sim(list(ma = 0.8), n = 100)
plot(testMA1, 
     ylab = "", main = "MA(1)")
testMA2 <- arima.sim(list(ma = c(-1, +0.8)), n = 100)
plot(testMA2, 
     ylab = "", main = "MA(2)")

```

Code for MA smoothing:
```{r MAsmooth}
plot(elecequip, main = "Electrical equipment manufactured in the Euro area", ylab = "Index, 2005 = 100")
lines(ma(elecequip, order = 3), col = 'red')
legend("topleft", cex = 0.75,legend=c("Level","3 month MA"), col=c("black", "red"), lty = c(1,1))

```

It's the Dow Jones again!  Remember we said DJ isn't stationary so we take differences?  Here's how the ARIMA process models them:

```{r arima dj}

djarima <- auto.arima(dj, seasonal = FALSE)
djarima
diffdjarima<- auto.arima(diff(dj), seasonal = FALSE)
diffdjarima

```

What about forecasting? Does this method help with forecasting the Dow?

```{r}
par(mfrow = c(1, 2))

plot(forecast(djarima,h=10),include=80)
plot(rwf(dj, drift = TRUE))


plot(forecast(diffdjarima,h=10),include=80)
plot(rwf(diff(dj), drift = TRUE))


par(mfrow = c(1, 1))

```

How pictures can help with your auto.arima check

```{r}
tsdisplay(testAR2, main = "What is this?")


tsdisplay(testMA2, main = "What about this one?")
```

Example:

Basic EDA:

```{r}
plot(elecequip)
Acf(elecequip)
```

Hmmmm looks like seasonal adjustment might help:

```{r}
eeadj <- seasadj(stl(elecequip, s.window="periodic"))
plot(elecequip)
lines(eeadj, col = "blue")
tsdisplay(diff(eeadj),main="")

```

ACF looks like there might be 3 relevant lags.


```{r}
fit <- Arima(eeadj, order=c(3,0,0))
summary(fit)
```

Maybe we need an ARIMA model, not an ARMA model

```{r}
fit <- Arima(eeadj, order=c(3,1,0))
summary(fit)
```

AICc went down.  What about a MA term?  There's some spikes in the PACF...

```{r}
fit <- Arima(eeadj, order=c(3,1,3))
summary(fit)
```

AICc went up. Wait, the MA term depends on the spikes in the ACF.  Maybe 3 is the wrong number

```{r}
fit <- Arima(eeadj, order=c(3,1,1))
summary(fit)
```

What does the auto.arima say?

```{r}
fit <- auto.arima(eeadj, seasonal = FALSE)
summary(fit)
```

Okay, let's go with this one.  How do the residuals look?

```{r}
Acf(residuals(fit))
```

Picture looks good but what about the statistics?

```{r}
Box.test(residuals(fit), lag=24, fitdf=4, type="Ljung")
```
What's with the fitdf = 4?

When testing the residuals of an estimated ARIMA model, the degrees of freedom need to be adjusted to reflect the parameter estimation. For example, for an ARIMA(p,0,q) model, the degrees of freedom should be set to p+q.


And more pictures?

```{r}
plot(forecast(fit))
```

Airpassenger seasonal example:

```{r}
tsdisplay(AirPassengers)

AP <- auto.arima(AirPassengers)
AP
APS <- auto.arima(AirPassengers, seasonal = FALSE)
APS

```
Wow, that's weird. But the AICc says that the seasonal model fits best.  Let's check out the residuals

```{r}

Acf(residuals(APS))

Acf(residuals(AP))

```

So we forecast:

```{r}
Box.test(residuals(AP), lag=24, fitdf=0, type="Ljung")

```
This says we didn't get it quite right.  Maybe we needed to do more work first.  Seasonally difference the data?  Some other transformation?



More elaborate example from the Fpp book


```{r}
tsdisplay(euretail)
```

***

```{r}
tsdisplay(diff(euretail,4))
```

***

```{r}
tsdisplay(diff(diff(euretail,4)))
```

***

```{r}
fit <- Arima(euretail, order=c(0,1,1), seasonal=c(0,1,1))
tsdisplay(residuals(fit))
```

***

```{r}
fit3 <- Arima(euretail, order=c(0,1,3), seasonal=c(0,1,1))
res <- residuals(fit3)
tsdisplay(res)
```

***

```{r}
Box.test(res, lag=16, fitdf=4, type="Ljung")
```

***

```{r}
auto.arima(euretail)
```

***

```{r}
auto.arima(euretail, stepwise=FALSE, approximation=FALSE)
```

***

```{r}
plot(forecast(fit3, h=12))
```

