===
Week2_class.Rmd
Requires: quantmod, fma, forecast packages, fpp
===



```{r}
library(quantmod)
getSymbols("DJIA", src = "FRED",auto.assign=TRUE)  #-- takes ticker names and pulls data from FRED
class(DJIA)
head(DJIA)
end(DJIA)


```
Now we want to reformat this daily data - we still want a time series but we want to get rid of weekends and holidays

```{r}
temp <- ts(na.omit(DJIA))
mydj <- ts(temp[1:length(temp)], start = 1)
class(mydj)
head(mydj)
tail(mydj)
plot(mydj, main = "Dow Jones Industrial average \n 30 Oct 2008 - 26 Oct 2018", ylab = "Index")

```

There is too much going on here - we will limit the series to the last year which is about 250 trading days

```{r}

mydj.short <- window(mydj, start = length(mydj)-250, end = length(mydj)) #note the indexes - you can change this if you want

plot(mydj.short, main = "Dow Jones Industrial average \n 250 trading days ending 29 Aug 2017",  ylab = "Index")

#create the change in the DJ 
diff_mydj <- diff(mydj.short)
plot(diff_mydj, main = "Change in daily Dow Jones Industrial average \n 250 trading days ending 26 Oct 2018", ylab = "Change in index")
abline(a=0,b=0, col = "red")


```


Okay, data munging and initial charting done.  Is this series white noise?  We'll look at a picture and calculate a Ljung Box statistic which requires that we choose lag length (h) to test. 

How do you choose h? FPP suggests h=10 for non-seasonal data, and h=2*m for seasonal data.  Hyndman describes his reasoning here: 
http://robjhyndman.com/hyndsight/ljung-box-test/

There are other arguments for other choices.  For example, see
http://stats.stackexchange.com/questions/6455/how-many-lags-to-use-in-the-ljung-box-test-of-a-time-series

```{r}
acf(diff_mydj)

#now calculate the test statistic
Box.test(diff_mydj, lag=10, fitdf=0, type="Lj")


```
So, is this white noise?  Remember we are dealing with a null hypothesis of no serial correlation.  If we get a pvalue greater than our threshhold (say 5% for 95% confidence), then we fail to reject the null.  You'll need to re-learn how to get your head around that.

Side note:  there is another test statistic that you will often see used - the Box Pierce statistic.  The calculation is very similar to the Ljung Box but it doesn't do as well with larger values of h.  The Box.test command in R defaults to the Box Pierce statistic if you don't specify type.

```{r}
#now calculate the test statistic
Box.test(diff_mydj, lag=10, fitdf=0)
```

You can see for this example, the statistic (which has the same Chi-square distribution to test the same hypothesis) is very close and the conclusion is the same. 


What do the residuals look like? 



Now: more on time series decomposition from week 1.  The R commands are very simple.  Here are the ones used to generate the charts in the slides:

```{r}
plot(decompose(hsales))
plot(stl(hsales,s.window="periodic"))
```

Similar but different calculations. 
You can pull the pieces out of the decomposition process

```{r}
hsales.decomp <- decompose(hsales)
plot(hsales.decomp$seasonal, main = "Home sales seasonal pattern")
plot(hsales.decomp$trend, main = "Home sales trend pattern")
```

Now we will decompose the unemployment rate
Get data from FRED.  Note the convoluted way that the time series item is created.  The stl package requires a particular shape to the time series that doesn't always result from the ts command. Note the differences in the dimensions and the description in the "Environment" window.

I have searched high and low for a more straightforward way to deal with this issue and welcome better code than I have written (hint: bonus points possible!)

```{r}
urates <- c("UNRATENSA", "UNRATE")
getSymbols(urates, src="FRED")  #-- takes ticker names and pulls data from FRED
start(UNRATENSA)
end(UNRATENSA)
class(UNRATENSA)
#suggestions for improving the 2 step process of creating a time series element with the correct dimensions are welcome

#first series
temp <- ts(UNRATENSA)
dim(temp)
ur.nsa <- ts(temp[1:823], start = c(1948,1), end = c(1948, length(UNRATENSA)),frequency = 12)
dim(ur.nsa)

#second series
temp <- ts(UNRATE)
ur<- ts(temp, start = c(1948,1), end = c(1948, length(UNRATE)), frequency = 12)

plot(ur, main = "Full unemployment rate from FRED")

```

We are only going to work with the last 10 years. Note one retains the seansonailty and other other is "deseasonized (adjusted)"

```{r}
ur.nsa.short <- window(ur.nsa, start = 2006)
ur.short <- window(ur, start = 2006)

plot(ur.nsa.short, ylab = "Percent", main = "Civilian unemployment rate")
lines(ur.short, col = "blue")
legend("topleft", legend=c("Unadjusted","Adjusted"), col=c("black","blue"), lty = 1)
```

Now to compare the decomposition.  Note: you cannot pass a title to the plot from decompose command without writing a convoluted wrapper function which is too painful for me. 

```{r}
plot(decompose(ur.nsa.short))
plot(stl(ur.nsa.short, s.window = "periodic"))

```

Let's compare the seasonal calculations:


```{r}
ur.nsa.decomp <- decompose(ur.nsa.short)
ur.nsa.stl <- stl(ur.nsa.short, s.window = "periodic")

plot(ur.nsa.decomp$seasonal, main = "Unemployment seasonal pattern")
lines(ur.nsa.stl$time.series[,1], col = "blue")
```

We can also use the seasadj command to take either of these seasonal factors and adjust the unadjusted rate.  How does that compare to the published rate?
```{r}
#decomposition first, then adjustment
fit <- stl(ur.nsa.short,s.window=15)
ur.nsa.short.sa <- seasadj(fit)

#let's see the results

plot(ur.nsa.short, col="gray", ylab = "Percent", main = "Adjusting unemployment")
fit <- stl(ur.nsa.short,s.window=15)
ur.nsa.short.sa <- seasadj(fit)
lines(ur.nsa.short.sa, col="red")
lines(ur.short, col = "blue")
legend("topright", legend=c("Unadjusted","R-adjusted","Published adj"), col=c("grey", "red", "blue"), lty = c(1,1,1))
```

Other ways to draw pictures of seasonal components:

Seasonal plots:

Data plotted against the individual “seasons” in
which the data were observed. For most of our examples,a
“season” is a month.)
Something like a time plot except that the data
from each season are overlapped.
Enables the underlying seasonal pattern to be
seen more clearly, and also allows any
substantial departures from the seasonal
pattern to be easily identified.


Another Example with the Forecast Package

```{r}
library(forecast)
seasonplot(window(hsales,start = 1985), year.labels = TRUE, col = 1:10, main = "Seasonal plots of home sales")
```

Seasonal sub series plots:
Data for each season collected together in time
plot as separate time series.
Enables the underlying seasonal pattern to be
seen clearly, and changes in seasonality over
time to be visualized.
```{r}
library(fpp)
beer <- window(ausbeer,start=1992) %>%
  plot(main = "Australian beer sales") %>%
  seasonplot(year.labels=TRUE, col = 1:14) %>%
  monthplot

```

