---
title: "Week 1-Zoo"
output: html_document
---

# Time Series using ZOO
#LOAD LIBRARIES
#you may need to install packages, if this is the first time you use them. Select Packages > Install Packages in R/RStudio) 


setwd("/Users/mylesgartland/OneDrive - Rockhurst University/Courses/Forecasting/TimeSeries/time_series/Week 1")


```{r}
library(tseries)
library(zoo)
#import dataset into a dataframe
setwd("/Users/mylesgartland/OneDrive - Rockhurst University/Courses/Forecasting/Time Series/time_series/Week 1")
cisco = read.table('Data/cisco_00-10.csv', header=T, sep=',')
# create time series for cisco prices
ciscots = zoo(cisco$Price, as.Date(as.character(cisco$Date), format = "%m/%d/%y"))
```


```{r}
# sort data in chronological order
# set variable Date as time/date variable
cisco$Date=as.Date(as.character(cisco$Date), format = "%m/%d/%y")
cisco=cisco[order(cisco$Date),]
```
#Creating new variables 
```{r}                     
# create lagged series using function lag(tsobject, k==1);
pricelag = lag(ciscots, k=-1);
head(pricelag)
# diff = p_t - p_(t-1);
pricedif = diff(ciscots);

#compute simple returns ret = (p_t-p_(t-1))/p_(t-1)
ret=(ciscots-pricelag)/pricelag
```

```{r}
# sort data in chronological order
# set variable Date as time/date variable
cisco$Date=as.Date(as.character(cisco$Date), format = "%m/%d/%y")
cisco=cisco[order(cisco$Date),]
```
#Creating new variables 


```{r}                     
# create lagged series using function lag(tsobject, k==1);
pricelag = lag(ciscots, k=-1);
head(pricelag)
# diff = p_t - p_(t-1);
pricedif = diff(ciscots);

#compute simple returns ret = (p_t-p_(t-1))/p_(t-1)
ret=(ciscots-pricelag)/pricelag
#Example of data analysis for cisco dataset
```


```{r}    
#DEFINE LOG RETURNS
#rts is a time series object since it is created from a TS object
rts = diff(log(ciscots))

#to retrieve numerical values from time series use coredata()
# rt is a numerical vector (no date information)
rt=coredata(rts)
#print first 6 values
head(rt)
```

# LOAD LIBRARIES
# Load fBasics packages into current session
# To install the package the first time, 
# select Tools from top Menu and select Install Packages 

```{r}
library(fBasics)
                     
# COMPUTE SUMMARY STATISTICS
basicStats(rt) 
           
# CREATE HISTOGRAM 
# OPTIONAL creates 2 by 2 display for 4 plots 
# par(mfcol=c(2,2)) 
hist(rt, xlab="Cisco log returns", prob=TRUE, main="Histogram") 
# add approximating normal density curve 
xfit<-seq(min(rt),max(rt),length=40) 
yfit<-dnorm(xfit,mean=mean(rt),sd=sd(rt)) 
lines(xfit, yfit, col="blue", lwd=2) 
```

```{r}
# CREATE NORMAL PROBABILITY PLOT 
qqnorm(rt) 
qqline(rt, col = 2) 
plot(rt)
                 
```

```{r}
# NORMALITY TESTS 
# Perform Jarque-Bera normality test. 
normalTest(rts,method=c("jb")) 
```

```{r}
# COMPUTE ACF AND PLOT CORRELOGRAM 
#prints acf values to console 
acf(rt, plot=F) 
#plot acf values on graph (correlogram) 
acf(rt, plot=T) 
```

```{r}
# COMPUTE LJUNG-BOX TEST FOR WHITE NOISE (NO AUTOCORRELATION) 
# to Lag 6 
Box.test(rts,lag=6,type='Ljung') 
# to Lag 12 
Box.test(rts,lag=12,type='Ljung') 
                     
```


