===
Week3-4_class.Rmd
Requires: fpp, forecast packages
===
 
Most of this code is adapted from the code available with the FPP book
First let's look at how to implement simple exponential smoothing (SES)


```{r}
require(fpp)
# Simple exponential smoothing - code from the FPP book 

oildata <- window(oil, start = 1996, end = 2007)
plot(oildata, ylab = "Oil (millions of tonnes)", xlab = "Year")

fit1 <- ses(oildata, alpha = 0.2, initial = "simple", h = 3)
fit2 <- ses(oildata, alpha = 0.6, initial = "simple", h = 3)
fit3 <- ses(oildata, h = 3) #note no value for alpha and no initial value.  The algorithm will calculate alpha as part of the optimization. 

#what kind of objects are these?
class(fit3)

#what is in them
summary(fit3)

```
Notice:  we used the SES command and it executed the ETS(A,N,N) command.  ses is just a wrapper for ETS.  

Also notice the estimate for alpha.  That came from the optimization. So did the initial state number (447.48)

Let's look at some pictures.


```{r}
plot(fit1, PI=FALSE, ylab="Oil (millions of tonnes)",
     xlab="Year", main="", fcol="white", type="o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")

legend("topleft",lty=1, col=c(1,"blue","red","green"), cex = 0.75, 
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.89)),pch=1)

```


This just covers our existing data - we've fitted the data with several options for the smoothing parameter.  Notice how flat things get when alpha is really small.  

Now let's try forecasting - we're fitting these smoothing methods to data through 2007 but our data really goes through 2010
```{r}
fit1 <- ses(oildata, alpha = 0.2, initial = "simple", h = 3)
fit2 <- ses(oildata, alpha = 0.6, initial = "simple", h = 3)
fit3 <- ses(oildata, h = 3)

plot(fit1, PI=FALSE, ylim = c(400, 600), ylab="Oil (millions of tonnes)",
xlab="Year", main="", fcol="white", type="o")
#this just plots the data through 2007 because that's what we ran the model over
#we have more data but have to plot it seperately if we want to compare the forecast with the actual 
# add these lines for putting the actual data past 2007
test.oil <- window(oil, start = 2007)
lines(test.oil, col = "black", type = "o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, cex =0.75, col=c(1,"blue","red","green"),
c("data", expression(alpha == 0.2), expression(alpha == 0.6),
expression(alpha == 0.89)),pch=1)
```
Next set of methods: Holt exponential smoothing.

```{r}
fith <- holt(oildata, alpha=0.8, beta=0.2, initial="simple", h=3) 
class(fith)
summary(fith)
oildata[1]
```
Notice that it shows the parameter settings we gave it. We now have a setting for the effect of the trend.  What's the initial value?  It's the first value of the oildata series - because we set the initialization routine = "simple"

Adding these lines to our chart.  What is the effect of adding trend? 
```{r}
plot(fith, PI=FALSE, ylim = c(400, 600))
lines(fitted(fith), type = "o", col="purple") 
lines(fith$mean, col="purple", type="o") 
lines(test.oil, col = "black", type = "o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, cex = 0.75, col=c(1,"blue","red","green", "purple"),
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.89),"Holt"),pch=1)
```

That was the additive trend model - let's make it muliplicative.  
```{r}
fith2 <- holt(oildata,alpha=0.8,beta=0.2,initial="simple",exponential=TRUE,h=3) 
summary(fith2)
```
We specified alpha and beta and told it to start with the initial value and the initial trend.

Now add that line to the chart:
```{r}
plot(fith2, PI=FALSE, ylim = c(400, 600))
lines(fitted(fith2), col="brown", type = "o")
lines(fith2$mean, col="brown", type="o")
legend("topleft", cex =0.75, lty=1, col=c(1,"blue","red","green", "purple", "brown"),
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.89),"Holt","Exponential"),pch=1)


```

Let's compare the simple exponential smoothing with the Holt model 

```{r}
air <- window(ausair,start=1990,end=2004)
fit0 <- ses(air, alpha = 0.8, beta = 0.2,initial = "simple", h=5)
fit1 <- holt(air, alpha=0.8, beta=0.2, initial="simple", h=5) 
fit2 <- holt(air,alpha=0.8,beta=0.2,initial="simple",exponential=TRUE,h=5) 

fit3 <- holt(air, alpha=0.8, beta=0.2, damped=TRUE, h=5)
fit4 <- holt(air, alpha=0.8, beta=0.2, damped=TRUE, exponential = TRUE, h=5) 


# Results for first model:
fit1$model$state
fitted(fit1)
fit1$mean
```
The forecast object has the fitted and the forecast values as well as the values for the trend and the level.

You can plot those pieces:
```{r}
plot(fit2$model$state)

```


```{r}
plot(fit2, type="o", ylab="Air passengers in Australia (millions)", 
     xlab="Year", fcol="white", PI=FALSE)
lines(fitted(fit0), col="brown") 
lines(fitted(fit1), col="blue") 
lines(fitted(fit2), col="red")
lines(fitted(fit3), col="green")
lines(fitted(fit4), col="purple")
lines(fit0$mean, col="brown", type="o") 
lines(fit1$mean, col="blue", type="o") 
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
lines(fit4$mean, col="purple", type="o")

legend("topleft", lty=1, cex=0.6, col=c("black","brown","blue","red","green", "purple"), 
       c("Data","SES", "Holt","Exponent","Additive damped trend", "Mult damped trend"))
```

So we can see how the forecast differs across various methods.  What is the effect of changing the parameters for any single method?
```{r}


livestock2 <- window(livestock,start=1970,end=2000)
fit1 <- holt(livestock2,alpha = 0.8, beta = .1)
fit2 <- holt(livestock2, alpha = .6, beta = .1)
fit3 <- holt(livestock2,alpha = .4, beta = .1)



plot(fit3, type="o", ylab="Livestock, sheep in Asia (millions)", ylim = c(250, 475),
     flwd=1, PI=FALSE, xlab = "beta = 0.1")
lines(window(livestock,start=2001),type="o")
lines(fit1$mean,col=2)
lines(fit2$mean,col=3)

legend("topleft", cex = 0.75, lty=1, pch=1, col=1:4,
       c("Data","Alpha = 0.8","Alpha = 0.4","Alpha = 0.2"))

```

Trend parameter:
```{r}

livestock2 <- window(livestock,start=1970,end=2000)
fit1 <- holt(livestock2,alpha = 0.8, beta = .1)
fit2 <- holt(livestock2, alpha = .8, beta = .5)
fit3 <- holt(livestock2,alpha = .8, beta = .7)



plot(fit3, type="o", ylab="Livestock, sheep in Asia (millions)", ylim = c(250, 475),
     flwd=1, PI=FALSE, xlab = "alpha = .8")
lines(window(livestock,start=2001),type="o")
lines(fit1$mean,col=2)
lines(fit2$mean,col=3)

legend("topleft", cex = 0.75, lty=1, pch=1, col=1:4,
       c("Data","Beta = 0.1","Beta = 0.5","Beta = 0.7"))

```



HW Seasonal models
Let's look at Tourists to Australia
```{r}
aust <- window(austourists,start=2005)
fit1hw <- hw(aust,seasonal="additive")
fit2hw <- hw(aust,seasonal="multiplicative")
fit3hw <- hw(aust,seasonal="additive", damped = TRUE)

plot(fit2hw,ylab="International visitor night in Australia (millions)",
     PI=FALSE, type="o", fcol="white", xlab="Year")
lines(fitted(fit1hw), col="red", lty=2)
lines(fitted(fit2hw), col="green", lty=2)
lines(fit1hw$mean, type="o", col="red")
lines(fit2hw$mean, type="o", col="green")
legend("topleft",lty=1, cex = 0.75, pch=1, col=1:3, 
  c("data","Holt Winters' Additive","Holt Winters' Multiplicative"))
```
We can pull the different pieces out of the forecast object and compare them across specifications.  What's the difference between additive seasonal and multiplicative seasonal?

```{r}
states <- cbind(fit1hw$model$states[,1:3],fit2hw$model$states[,1:3])
colnames(states) <- c("level","slope","seasonal","level","slope","seasonal")
plot(states, xlab="Year", main = "Comparing Additive and Multiplicative Seasonals")
```

Let's revisit the oil data - we originally used the SES model on it which is equivalent to ETS(A,N,N)

```{r}
oildata <- window(oil, start = 1996, end = 2007)
fita <- ets(oildata, model = "ANN")
plot(forecast(fita, h=3), ylab="Oil (millions of tonnes)")
summary(fita)
ls(fita)  #list names of the objects in the specified environment
fita$par

```

What if we used the ETS(M,N,N) model?

```{r}
oildata <- window(oil, start = 1996, end = 2007)
fitm <- ets(oildata, model = "MNN")
plot(forecast(fitm, h=3), ylab="Oil (millions of tonnes)")
summary(fitm)
ls(fitm)  #list names of the objects in the specified environment
fitm$par

```
But are either of these the right model for these data?  What if we let the ETS code decide?

```{r}
oildata <- window(oil, start = 1996, end = 2007)
fit <- ets(oildata)
summary(fit)
plot(fit)
plot(forecast(fit, h = 8), ylab = "Oil (millions of tonnes)")

```


The model chose additive.  Compare the AIC for the two we examined.

```{r}
fita$aic
fitm$aic
```
Yup - the AIC for the additive model is lower so it is the "better" model 

How about a more interesting example?
Remember the Australian visitors data?
```{r}
aust <- window(austourists,start=2005)
plot(aust, main = "International visitor night in Australia", ylab = "Millions")
```
Very strongly seasonal.  Probably not a good candidate for the ETS(A,N,N) model.  Let's see:

```{r}
fit <- ets(aust)
summary(fit)

plot(fit)

plot(forecast(fit, h = 8), ylab = "Millions")


```
The winner is multiplicative errors, no trend, no seasonality.


What about our ubiquitous air passengers data?
```{r}
ap <- window(AirPassengers, start = 1949, end = 1958)
fit <- ets(ap)
summary(fit)

plot(fit)

plot(forecast(fit, h = 8), ylab = "Millions")


```

Same thing. 