---
title: "dynamic regression"
output: html_document
---
title:: Applied forecasting for business and economics"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE,
  dev.args=list(bg=grey(0.9), pointsize=11))
library(fpp2)
```

# Regression with ARIMA errors

## US personal consumption and income
```{r usconsump, fig.height=6, fig.width=8.5}
autoplot(uschange, facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("Quarterly changes in US consumption and personal income")
```

## US personal consumption and income


```{r, fig.height=6, fig.width=8.5}
ggplot(aes(x=Income,y=Consumption), data=as.data.frame(uschange)) +
  geom_point() +
  ggtitle("Quarterly changes in US consumption and personal income")
```

## US personal consumption and income

  * No need for transformations or further differencing.
  *  Increase in income does not necessarily translate into instant increase in consumption (e.g., after the loss of a job, it may take a few months for expenses to be reduced to allow for the new circumstances). We will ignore this for now.

## US personal consumption and income
\fontsize{11}{14}\sf

```{r usconsump2, echo=TRUE, fig.height=3}
(fit <- auto.arima(uschange[,1], xreg=uschange[,2]))
```


## US personal consumption and income
\fontsize{13}{15}\sf

```{r , echo=TRUE, fig.height=3.7}
ggtsdisplay(residuals(fit, type='regression'),
  main="Regression errors")
```

## US personal consumption and income
\fontsize{13}{15}\sf

```{r , echo=TRUE, fig.height=3.7}
ggtsdisplay(residuals(fit, type='response'),
  main="ARIMA errors")
```

## US personal consumption and income

```{r , echo=TRUE, fig.height=3.7}
checkresiduals(fit, test=FALSE)
```

## US personal consumption and income
\fontsize{10}{13}\sf

```{r , echo=TRUE, fig.height=3.7}
checkresiduals(fit, plot=FALSE)
```

## US personal consumption and income
\fontsize{9}{12}\sf

```{r usconsump3, echo=TRUE, fig.height=3.}
fcast <- forecast(fit,
  xreg=rep(mean(uschange[,2]),8), h=8)
autoplot(fcast) + xlab("Year") +
  ylab("Percentage change") +
  ggtitle("Forecasts from regression with ARIMA(1,0,2) errors")
```

## Forecasting

  * To forecast a regression model with ARIMA errors, we need to forecast the
regression part of the model and the ARIMA part of the model and combine the
results.
  * Some predictors are known into the future (e.g., time, dummies).
  * Separate forecasting models may be needed for other predictors.
  * Forecast intervals ignore the uncertainty in forecasting the predictors.

## Daily electricity demand
\fontsize{12}{13}\sf

Model daily electricity demand as a function of temperature using quadratic regression with ARMA errors.

\fontsize{10}{12}\sf

```{r, echo=TRUE, fig.height=3.6}
qplot(elecdaily[,"Temperature"], elecdaily[,"Demand"]) +
  xlab("Temperature") + ylab("Demand")
```

## Daily electricity demand
\fontsize{12}{13}\sf

```{r, echo=TRUE, fig.height=4}
autoplot(elecdaily, facets = TRUE)
```

## Daily electricity demand
\fontsize{9}{10}\sf

```{r, echo=TRUE}
xreg <- cbind(MaxTemp = elecdaily[, "Temperature"],
              MaxTempSq = elecdaily[, "Temperature"]^2,
              Workday = elecdaily[, "WorkDay"])
fit <- auto.arima(elecdaily[, "Demand"], xreg = xreg)
summary(fit)
checkresiduals(fit)
```

## Daily electricity demand
\fontsize{9}{10}\sf

```{r, echo=FALSE}
checkresiduals(fit, plot=FALSE)
```

## Daily electricity demand
\fontsize{10}{13}\sf

```{r, echo=TRUE}
# Forecast one day ahead
forecast(fit, xreg = cbind(26, 26^2, 1))
```

## Daily electricity demand
\fontsize{10}{13}\sf
```{r echo=TRUE}
fcast <- forecast(fit,
  xreg = cbind(rep(26,14), rep(26^2,14),
    c(0,1,0,0,1,1,1,1,1,0,0,1,1,1)))
autoplot(fcast) + ylab("Electicity demand (GW)")
```


# Stochastic and deterministic trends

## Stochastic \& deterministic trends

## International visitors
```{r}
autoplot(austa) + xlab("Year") +
  ylab("millions of people") +
  ggtitle("Total annual international visitors to Australia")
```
## International visitors
\fontsize{10}{10}\sf
**Deterministic trend**
```{r, echo=TRUE}
trend <- seq_along(austa)
(fit1 <- auto.arima(austa, d=0, xreg=trend))
```
\pause\vspace*{-0.1cm}
```{r austaparams, echo=FALSE}
phi1 <- coef(fit1)['ar1']
phi2 <- coef(fit1)['ar2']
intercept <- coef(fit1)['intercept']
slope <- coef(fit1)['xreg']
sigma2 <- fit1$sigma2
```
###

## International visitors
\fontsize{10}{10}\sf

**Stochastic trend**

```{r, echo=TRUE}
(fit2 <- auto.arima(austa,d=1))
```
\pause\vspace*{-0.1cm}


```{r austaparams2, cache=TRUE, echo=FALSE}
drift <- coef(fit2)['drift']
theta1 <- coef(fit2)['ma1']
sigma2 <- fit2$sigma2
```

## International visitors

```{r, fig.height=2.2}
autoplot(forecast(fit1, xreg=length(austa) + 1:10)) +
  xlab("Year") + ylab("") +
  ggtitle("Forecasts from linear trend with AR(2) error")
```

```{r, fig.height=2.2}
autoplot(forecast(fit2)) +
  xlab("Year") + ylab("") +
  ggtitle("Forecasts from ARIMA(0,1,1) with drift")
```

## International visitors

```{r echo=FALSE, fig.height=5.5, fig.width=8.5}
fc1 <- forecast(fit1,
  xreg = data.frame(trend = length(austa) + 1:10))
fc2 <- forecast(fit2, h=10)
autoplot(austa) +
  autolayer(fc2, series="Stochastic trend") +
  autolayer(fc1, series="Deterministic trend") +
  ggtitle("Forecasts from trend models") +
  xlab("Year") + ylab("Visitors to Australia (millions)") +
  guides(colour=guide_legend(title="Forecast"))
```

## Forecasting with trend

  * Point forecasts are almost identical, but prediction intervals differ.
  * Stochastic trends have much wider prediction intervals because the errors are non-stationary.
  * Be careful of forecasting with deterministic trends too far ahead.

# Dynamic harmonic regression

## Dynamic harmonic regression

 **Combine Fourier terms with ARIMA errors**

### Advantages
   * it allows any length seasonality;
   * for data with more than one seasonal period, you can include Fourier terms of different frequencies;
   * the seasonal pattern is smooth for small values of $K$ (but more wiggly seasonality can be handled by increasing $K$);
   * the short-term dynamics are easily handled with a simple ARMA error.

### Disadvantages
 * seasonality is assumed to be fixed


## Eating-out expenditure

```{r cafe, echo=TRUE, fig.height=4.6, fig.width=8}
cafe04 <- window(auscafe, start=2004)
autoplot(cafe04)
cafefit <- function(K)
{
  require(latex2exp)
  fit <- auto.arima(cafe04, xreg=fourier(cafe04, K=K),
                  seasonal = FALSE, lambda = 0)
  reg <- log(cafe04) - residuals(fit, type='regression')
  reg <- exp(reg - mean(reg) + mean(log(cafe04)))
  fc <- fit %>%
    forecast(xreg=fourier(cafe04, K=K, h=24))
  autoplot(cafe04, series="Data") +
    autolayer(fc) + ggtitle(TeX(paste(fc$method,"and $\\lambda = 0$"))) +
    autolayer(reg, series="Regression fit") +
    xlab(paste("K=",K,"   AICC=",round(fit$aicc,2))) +
    ylab("") + ylim(1.5,4.7)
}
```

## Eating-out expenditure

```{r cafe1, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(1)
```

## Eating-out expenditure

```{r cafe2, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(2)
```

## Eating-out expenditure

```{r cafe3, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(3)
```

## Eating-out expenditure

```{r cafe4, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(4)
```

## Eating-out expenditure

```{r cafe5, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(5)
```

## Eating-out expenditure

```{r cafe6, dependson='cafe', fig.height=5, echo=FALSE}
cafefit(6)
```

## Example: weekly gasoline products
\fontsize{8}{8}\sf

```{r gasmodel, echo=TRUE}
harmonics <- fourier(gasoline, K = 13)
(fit <- auto.arima(gasoline, xreg = harmonics, seasonal = FALSE))
```

## Example: weekly gasoline products
\fontsize{11}{12}\sf

```{r gasf, echo=TRUE, fig.height=3.5}
newharmonics <- fourier(gasoline, K = 13, h = 156)
fc <- forecast(fit, xreg = newharmonics)
autoplot(fc)
```

## 5-minute call centre volume

```{r calls, echo=TRUE, fig.height=4}
autoplot(calls)
```

## 5-minute call centre volume
\fontsize{8}{8}\sf

```{r callsmodel, echo=TRUE}
xreg <- fourier(calls, K = c(10,0))
(fit <- auto.arima(calls, xreg=xreg, seasonal=FALSE, stationary=TRUE))
```

## 5-minute call centre volume

```{r callsres, echo=TRUE}
checkresiduals(fit, test=FALSE)
```

## 5-minute call centre volume
\fontsize{10}{11}\sf

```{r callsf, echo=TRUE, fig.height=4}
fc <- forecast(fit, xreg = fourier(calls, c(10,0), 1690))
autoplot(fc)
```

# Lagged predictors

## Lagged predictors


## Example: Insurance quotes and TV adverts
```{r tvadvert}
autoplot(insurance, facets=TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("Insurance advertising and quotations")
```
## Example: Insurance quotes and TV adverts
\fontsize{10}{10}\sf
```{r, echo=TRUE}
Advert <- cbind(
    AdLag0 = insurance[,"TV.advert"],
    AdLag1 = lag(insurance[,"TV.advert"],-1),
    AdLag2 = lag(insurance[,"TV.advert"],-2),
    AdLag3 = lag(insurance[,"TV.advert"],-3)) %>%
  head(NROW(insurance))
# Restrict data so models use same fitting period
fit1 <- auto.arima(insurance[4 :40,1], xreg=Advert[4:40,1],
  stationary=TRUE)
fit2 <- auto.arima(insurance[4:40,1], xreg=Advert[4:40,1:2],
  stationary=TRUE)
fit3 <- auto.arima(insurance[4:40,1], xreg=Advert[4:40,1:3],
  stationary=TRUE)
fit4 <- auto.arima(insurance[4:40,1], xreg=Advert[4:40,1:4],
  stationary=TRUE)
c(fit1$aicc,fit2$aicc,fit3$aicc,fit4$aicc)
```
## Example: Insurance quotes and TV adverts
\fontsize{10}{10}\sf
```{r tvadvertagain, echo=TRUE}
(fit <- auto.arima(insurance[,1], xreg=Advert[,1:2],
  stationary=TRUE))
```
\pause
```{r tvadvertparam, echo=FALSE}
# Store coefficients
phi1 <- coef(fit)['ar1']
phi2 <- coef(fit)['ar2']
phi3 <- coef(fit)['ar3']
intercept <- coef(fit)['intercept']
gamma0 <- coef(fit)['AdLag0']
gamma1 <- coef(fit)['AdLag1']
```
###

## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf
```{r, echo=TRUE, fig.height=3.3}
fc <- forecast(fit, h=20,
  xreg=cbind(c(Advert[40,1],rep(10,19)), rep(10,20)))
autoplot(fc)
```
## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf
```{r, echo=TRUE, fig.height=3.3}
fc <- forecast(fit, h=20,
  xreg=cbind(c(Advert[40,1],rep(8,19)), rep(8,20)))
autoplot(fc)
```
## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf
```{r, echo=TRUE, fig.height=3.3}
fc <- forecast(fit, h=20,
  xreg=cbind(c(Advert[40,1],rep(6,19)), rep(6,20)))
autoplot(fc)
```
## Transfer function models
